Backward Population Synthesis: Mapping the Evolutionary History of Gravitational-Wave Progenitors

### 2023-02-10 src/tex/ms.tex
Rearrange paragraphs for case-specific configuration

In full: we propose a method to ``backward" model each GW event to its
progenitor state while allowing the hyperparameters to vary across their full
range of physical uncertainty following a two-stage process. First, we solve a
root finding problem to obtain guesses that are likely to produce the desired
system properties in the joint progenitor--hyperparameter space. We then sample
the posterior in the joint space that is induced by the DCO merger observations
using a Markov chain Monte Carlo (MCMC) algorithm that is initialized by the
roots found in the previous stage.  We demonstrate our algorithm in a realistic
setting by producing a posterior over progenitor properties and \texttt{COSMIC}
hyperparameters implied by the observation of GW150914.

@@

@@

For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
the critical mass ratio which determines whether Roche-overflow mass transfer
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$. \textbf{We follow the critical mass ratio prescription 
described in \cite{Neijssel2019} for donors which fill their Roche lobes during
 other phases of stellar evolution.} 
For systems which begin a common envelope 
with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
stars merge since the donor has likely not formed a strong core-envelope boundary 
\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
evolution proceed. In particular, we assume that stars lose mass through stellar winds 
in a metallicity-dependent 
fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
during both stable Roche-overflow mass transfer and common envelopes.
We further assume that the CO formation mechanism follows the delayed 
prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
and that natal kicks due to supernovae are reduced by the amount 
of mass which falls back onto the proto-CO
as described in \citet{Fryer2012}.}

>>>>

In full: we propose a method to ``backward" model each GW event to its
progenitor state while allowing the hyperparameters to vary across their full
range of physical uncertainty following a two-stage process. First, we solve a
root finding problem to obtain guesses that are likely to produce the desired
system properties in the joint progenitor--hyperparameter space. We then sample
the posterior in the joint space that is induced by the DCO merger observations
using a Markov chain Monte Carlo (MCMC) algorithm that is initialized by the
roots found in the previous stage.  We demonstrate our algorithm in a realistic
setting by producing a posterior over progenitor properties and \texttt{COSMIC}
hyperparameters implied by the observation of
dependencies:
  src/scripts/figure1_cornerPlot.py:
    - src/data/GWTC3_LVK_posterior.npz
    - src/data/GWTC3_mcmc_processed.hdf5

  src/scripts/figure2_varyAlphaFacc.py:
    - src/data/fig_2_dat.h5

  src/scripts/figure3_sanityCheck.py:
    - src/data/GWTC3_LVK_posterior.npz
    - src/data/GW150914_roots_raw.hdf5
    - src/data/GW150914_roots_processed.hdf5
    - src/data/GWTC3_mcmc_processed.hdf5

  src/scripts/figure4_population.py:
    - src/data/GWTC3_mcmc_processed.hdf5

datasets:
  10.5281/zenodo.6624911:
    destination: src/data
    contents:
      GW150914_roots_processed.hdf5:
      GW150914_roots_raw.hdf5:
      GWTC3_LVK_posterior.npz:
      GWTC3_mcmc_processed.hdf5:
      fig_2_dat.h5:
 GW150914.

@@

\textbf{We follow the critical mass ratio prescription 
described in \cite{Neijssel2019} for donors which fill their Roche lobes during
 other phases of stellar evolution.} 
For systems which begin a common envelope 
with a Hertzsprung-gap donor, we always assume the pessimistic outcome that the two 
stars merge since the donor has likely not formed a strong core-envelope boundary 
\citep{Ivanova2004, Belczynski2008}. \textbf{We fix other assumptions for how stellar and binary 
evolution proceed. In particular, we assume that stars lose mass through stellar winds 
in a metallicity-dependent 
fashion according to \cite{Vink2001,Vink2005}. We assume that binaries are rapidly circularized
during both stable Roche-overflow mass transfer and common envelopes.
We further assume that the CO formation mechanism follows the delayed 
prescription of \citet{Fryer2012} which allows neutrino-driven supernova explosions to 
occur from instabilities which grow on timescales longer than $250\,\rm{ms}$, 
and that natal kicks due to supernovae are reduced by the amount 
of mass which falls back onto the proto-CO
as described in \citet{Fryer2012}.}

\subsection{\textbf{Case-specific configuration}}
\label{subsec:configuration}

@@

For hyperparameters $\bm{\lambda}$, we choose parameters that strongly affect
binary evolution in \texttt{COSMIC} for massive stars, including the common envelope
efficiency $\alpha$, the fraction of mass accreted onto the accretor during stable mass transfer $f_{\rm acc}$,
the critical mass ratio which determines whether Roche-overflow mass transfer
proceeds stably or produces a common envelope when the donor star is on the Hertzsprung gap
$q_{\rm crit,3}$, and the root-mean-square of the Maxwellian distribution
employed for CO natal kicks, $\sigma$. 
%Prior
